---
title: "Final Project - Background"
date: "2022-12-08"
output: html_document
editor_options: 
  chunk_output_type: inline
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#### Background
In parsing data sets to find one usable for the purposes of this project, our group settled upon using an extensive database of spotify songs (tracks) (232,725 entries across 18 features). The features include a number of acoustic characteristics (loudness, tempo, valence, time signature, etc.), a popularity metric (from 0 to 100), and other identifying features like genre and names of the artist and track. All of the members of our team are deeply interested in music and saw this data set, given its expansiveness, as a good platform to build our work upon.

The dataset can be found here: [Kaggle - Spotify Tracks DB](https://www.kaggle.com/datasets/zaheenhamidani/ultimate-spotify-tracks-db)<br /> 
With additional credit for work to grab the data: [Github - Spotify Data Project](https://github.com/tgel0/spotify-data)

After deliberation our groups settled upon question was... <br /><br />
**Can we use the quantifiable metrics or characteristics that defined a song to predict song popularity once it is released under the label of a specific genre?**<br /><br />
Our group settled on this question as popularity was the metric we saw as most likely deterministic based on the other features present in the selected dataset. Stepping back from the specific data at hand, as well, popularity is a metric that is pertinent both to businesses - be that independent artists or recording labels - and us as consumers - as popularity undeniably plays a role in what we consume; the relevance to both artist and consumer make a worthy subject for a our research question. 

**Music Characteristics:** <br />
Per the Spotify Developer API (linked [here](https://developer.spotify.com/documentation/web-api/reference/#/operations/get-audio-features)) a single track has 13 unique features which are quantifiable and contribute to the audio profile. These are detailed in the table below with descriptions and possible value ranges.
```{r, include=FALSE}
xx <- read.csv("./spotify.csv")
xx[xx=="?"] <- NA
feats <- xx[1:13,]
misc <- xx[14:18,]
```
```{r}
knitr::kable(feats, col.names = c('Characterstic', 'Description', 'Value Range'), caption = 'Table 1. Predictive features from API')
```
There are 5 additional features that are returned by a *GET* call to the Spotify Dev API that are useful in labeling and management of data throughout the analytics process but are not of predictive value. These are detailed in second table below. Not all of these misc. features are present in the kaggle dataset used, but were included here for complete context.
```{r}
knitr::kable(misc, col.names = c('Characterstic', 'Description', 'Value Range'), caption = 'Table 2. Misc. features from API')
```
 
**Previous Works** <br />
From our research there was one example of a pre-existing project that was targeting the same question as we were. The work in question was conducted by Matt Devor; operating based on the same 13 characteristics we also identified from the API, Devor performed a range of data analysis and built towards running a Linear Regression model to directly predict popularity. Devor's work is limited in the nature of the concluding model, but in lieu of that it does serve as an extensive reference for exploratory data analysis which our team looks to build upon in our work.

Reference: [Github - Predicting Spotify Song Popularity](https://github.com/MattD82/Predicting-Spotify-Song-Popularity)

**Our Approach** <br />
In order to address the landed upon question from above - exploring the predictive power of the acoustic characteristics of an audio track to determine a track's popularity - in a manner unique from existing work our team will use a random forest model; the model will perform regression over the continuous target variable *popularity*. Through exploratory analytics and further method decisions we seek to create a model with a sub 5% NMRSE as for use potential use in a real-world business context.



